A Transdiciplinary approach to developing a spectral numerical coastal-flow

### 2019-09-04 literature_review.Rmd
added to buoyancy/drag properties section and updated pdf and word outputs.

---
title: 'A Transdiciplinary approach to developing a spectral numerical coastal-flow
  model based on passive macroalgal rafting phenenomea to shed light on ambient and
  stochastic dispersal and connectivity of nearshore and offshore marine communities
  and 3D hyperspectral hydrodynamic modelling in and around the coast of South Africa'
author: "Ross Coppin"
output:
  pdf_document: 
  word_document:
    fig_caption: yes
    reference_docx: frontiers_template.docx
    toc: yes
---

@@

Waves are complex and various approaches, techniques and devices have been developed in order to measure waves effectively. One of the ways that has been used extensively used in the past but less so today, are visual estimates which can be used to characterise *significant swell height* (Hs) and *significant swell period* (Tp). Visual estimates use 15 - 20 of the most well defined, higher waves of a number of wave groups to characterise Hs and Tp. Although these parameters are useful, they do not accurately reflect the waves that are occurring in nature. Ocean waves are a combination of wind sea (short, irregular, locally generated waves) and swell (long, smooth waves, generated by distant storms) and so more parameters are needed to separate Hs and Tp driven by wind sea and swell, i.e. separate Hs and Tp parameters for wind sea and swell. However, even with separate parameters for the different types of waves would still not be enough to effectively characterise waves in the ocean. In order to characterise the detail and complexity of ocean waves a different approach must be used, know as the *spectral* approach. This approach is based on the idea that the sea surface can be characterised as the summation of a large number of harmonic wave components, known as the *random-phase/amplitude* model. The random-phase amplitude model is a summation of wave components on a discrete scale, however a continous scale is more relevant to characterising and measuring of waves in the ocean. The discrete spectrums are converted to continoius spectrums through various methods. Therefore the random-phase amplitude model is used to obtain a continious variance density spectrum.

@@

# Kelp drag properties

>>>>

---
title: A Transdiciplinary approach to developing a spectral numerical coastal-flow
  model based on passive macroalgal rafting phenenomea to shed light on ambient and
  stochastic dispersal and connectivity of nearshore and offshore marine communities
  and 3D hyperspectral hydrodynamic modelling in and around the coast of South Africa
author: "Ross Coppin"
output:
  pdf_document:
    toc: yes
  word_document:
    fig_caption: yes
    reference_docx: frontiers_template.docx
    toc: yes
---

@@

Waves are complex and various approaches, techniques and devices have been developed in order to measure waves effectively. One of the ways that has been used extensively used in the past but less so today, are visual estimates which can be used to characterise *significant swell height* (Hs) and *significant swell period* (Tp). Visual estimates use 15 - 20 of the most well defined, higher waves of a number of wave groups to characterise Hs and Tp. Although these parameters are useful, they do not accurately reflect the waves that are occurring in nature. Ocean waves are a combination of wind sea (short, irregular, locally generated waves) and swell (long, smooth waves, generated by distant storms) and so more parameters are needed to separate Hs and Tp driven by wind sea and swell, i.e. separate Hs and Tp parameters for wind sea and swell. However, even with separate parameters for the different types of waves would still not be enough to effectively characterise waves in the ocean. In order to characterise the detail and complexity of ocean waves a different approach must be used, know as the *spectral* approach. This approach is based on the idea that the sea surface can be characterised as the summation of a large number of harmonic wave components, known as the *random-phase/amplitude* model. The random-phase amplitude model is a summation of wave components on a discrete scale, however a continous scale is more relevant to characterising and measuring of waves in the ocean. The discrete spectrums are converted to continoius spectrums through various methods. Therefore, the random-phase amplitude model is used to obtain a *continious variance density spectrum* which is a statistical relevant measure for scientists and engineers.

@@

# Buoyancy/drag properties

In order to determine the fraction of the kelp raft submerged the *Buoyancy Force* (BF) must be calculated. A theoretical example using a object of a regular shape (see figure \ref{fig:BF_1}) will be used to demonstrate the theory to applied to this component of the study. First the fraction of the object submerged can be calculated, which in this case is represented by $x$ (see figure \ref{fig:BF_1}).

```{r BF_1, echo=FALSE, message=FALSE, warning=FALSE, fig.width = 3, fig.height = 3, fig.cap='\\label{fig:BF_1}Conceptual model of bouyancy characteristics.'}
BF_1 <- readPNG("figures/BF_1.png")
grid.raster(BF_1)
```

An object will only float if the object density ($\rho$) is less than or equal to the density of the liquid in which the object is submerged, i.e. $\rho_{object} < \rho_{liquid}$. The bouyancy force (*BF*) is the force of the displaced liquid pushing the object in an upwards direction which is equal to the weight of the displaced liquid, $BF = W_{displaced-liquid}$. The weight of the displaced liquid can be calculated by multiplying the mass of the liquid by the gravitational constant $g$ and so the weight of the displaced liquid can be calculated by: $$W_{displaced-liquid} = mg$$ The equation is expanded upon by including the calculation for mass, $m = \rho V$, and so now the weight of the displaced liquid can be calculated as: $$mg = \rho_{liquid} V g$$ where $\rho$ represents the density of the liquid and $V$ represents the portion of the object submerged. Since the object is floating, the BF can also be represented by the weight of the object and as mentioned previously the weight of the displaced liquid can also be represented by $mg$, which in turn can be written as $W_{displaced-liquid} = W_{object}$. Now through substituting the calculations for weight, volume and the dimensions of the object, the equation can be simplied to: $$x= h\displaystyle\frac{\rho_o}{\rho_l}$$ where $h$ is the overall height of the object, $\rho_o$ represents the density of the object and $\rho_l$ represents the density of the liquid. It is important to note that this equation is only applicable to objects that are floating i.e. $\rho_{object} < \rho_{liquid}$.
